Analyzing the influence of Selection on Genetic Programming's Generalization ability in Symbolic Regression

### 2022-06-26 docs/rmd/presentation/slides.Rmd
addded some tables

---
title: "Analyzing the influence of Selection on Genetic Programming's Generalization ability in Symbolic Regression"
subtitle: "A comparison of epsilon-lexicase Selection and Tournament Selection"
author: "Roman Hoehn, B.Sc. Wirtschaftspaedagogik"
date: "2022-06-29"
output:
  beamer_presentation:
    colortheme: "beaver"
    fonttheme: "structurebold"
    keep_tex: true
    fig_caption: yes
    number_sections: true
    slide_level: 2
toc: yes
# toc_depth: 3
bibliography: ../bib/bib/ref.bib
csl: ../bib/csl/harvard1.csl
---

@@

* Total dataset ($N=768$) is randomly split into a training and testing dataset (50:50)
* Train two models using GP with the training dataset only, one using tournament selection and the other $\epsilon$-lexicase selection
* For each generation: Select elite model and compute its fitness for the testing dataset

@@

## Genetic Programming Configuration

>>>>

---
title: "Analyzing the influence of Selection on Genetic Programming's Generalization ability in Symbolic Regression"
subtitle: "A comparison of epsilon-lexicase Selection and Tournament Selection"
author: "Roman Hoehn, B.Sc. Wirtschaftspaedagogik"
date: "2022-06-29"
output:
  beamer_presentation:
    colortheme: "beaver"
    fonttheme: "structurebold"
    #extra_dependencies: ["float"]
    keep_tex: true
    fig_caption: yes
    number_sections: true
    slide_level: 2
    #includes:
      #in_header: ./settings.tex
toc: yes
# toc_depth: 3
bibliography: ../bib/bib/ref.bib
csl: ../bib/csl/harvard1.csl
---

@@

* Total dataset ($N=768$) is randomly split into a training and testing dataset (50:50)
* Fitness metric: Mean squared Error (MSE)
* Train two models using GP with the training dataset only, one using tournament selection and the other $\epsilon$-lexicase selection
* For each generation: Select elite model and compute its fitness for the testing dataset

@@

## Evolutionary Parameters

knitr::kable(
  read.csv("../tables/csv/gp_parameters.csv"),
  digits=3,
  caption="Evolutionary Parameters"
) %>% row_spec(0,bold=TRUE ) %>% kableExtra::kable_styling(latex_options = "hold_position")
```

## Function Set

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
fig.pos="b"

knitr::kable(
  read.csv("../tables/csv/primitives.csv"),
  digits=3,
) 
```

## Terminal Set

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
fig.pos="t"

knitr::kable(
  read.csv("../tables/csv/terminals.csv"),
  digits=3,
) 
```

## Examples

Model evolved by tournament selection after 100 generations:

![](../plots/example_model.png)

### 2022-06-26 docs/rmd/presentation/slides.tex
addded some tables

\begin{frame}{epsilon-Lexicase Selection}
\protect\hypertarget{epsilon-lexicase-selection}{}
\begin{itemize}
\tightlist
\item
  Recent alternative: Lexicase Selection and it's variation
  \(\epsilon\)-lexicase selection
\item
  Idea: Selection method for uncompromising, continous-valued symbolic
  regression problems\footnote<.->{Helmuth, Spector and Matheson (2015),
    p.12}
\item
  Increases genetic diversity inside the population\footnote<.->{Helmuth,
    Spector and Matheson (2015), p.1}
\item
  Higher chance for ``specialist'' solutions to be selected since it is
  decided on a per case basis
\item
  Performance increases have been demosntrated in many benchmarking
  problems\footnote<.->{La Cava, Spector and Danai (2016), p.744-745}
\end{itemize}
\end{frame}

@@

\begin{frame}{Experiment}
\protect\hypertarget{experiment}{}
\begin{block}{Single run}
\protect\hypertarget{single-run}{}
\begin{itemize}
\tightlist
\item
  Total dataset (\(N=768\)) is randomly split into a training and
  testing dataset (50:50)
\item
  Train two models using GP with the training dataset only, one using
  tournament selection and the other \(\epsilon\)-lexicase selection
\item
  For each generation: Select elite model and compute its fitness for
  the testing dataset
\end{itemize}
\end{block}

@@

\begin{frame}{Genetic Programming Configuration}
\protect\hypertarget{genetic-programming-configuration}{}
\ldots{}
\end{frame}

>>>>

\begin{frame}{epsilon-Lexicase Selection}
\protect\hypertarget{epsilon-lexicase-selection}{}
\begin{itemize}
\tightlist
\item
  Recent alternative: Lexicase Selection and it's variation
  \(\epsilon\)-lexicase selection
\item
  Idea: Selection method for uncompromising, continous-valued symbolic
  regression problems\footnote<.->{Helmuth, Spector and Matheson (2015),
    p.12}
\item
  Increases genetic diversity inside the population\footnote<.->{Helmuth,
    Spector and Matheson (2015), p.1}
\item
  Higher chance for ``specialist'' solutions to be selected since it is
  decided on a per case basis
\item
  Performance increases have been demonstrated in many benchmarking
  problems\footnote<.->{La Cava, Spector and Danai (2016), p.744-745}
\end{itemize}
\end{frame}

@@

\begin{frame}{Experiment}
\protect\hypertarget{experiment}{}
\begin{block}{Single run}
\protect\hypertarget{single-run}{}
\begin{itemize}
\tightlist
\item
  Total dataset (\(N=768\)) is randomly split into a training and
  testing dataset (50:50)
\item
  Fitness metric: Mean squared Error (MSE)
\item
  Train two models using GP with the training dataset only, one using
  tournament selection and the other \(\epsilon\)-lexicase selection
\item
  For each generation: Select elite model and compute its fitness for
  the testing dataset
\end{itemize}
\end{block}

@@

\begin{frame}{Evolutionary Parameters}
\protect\hypertarget{evolutionary-parameters}{}
\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-2}Evolutionary Parameters}
\centering
\begin{tabular}[t]{l|l}
\hline
\textbf{Parameter} & \textbf{Value}\\
\hline
Population Size & 500\\
\hline
Number of Generations & 100\\
\hline
Mutation Rate & 20\%\\
\hline
Crossover Rate & 80\%\\
\hline
Tournament Size & 3\\
\hline
Epsilon selection & automatic\\
\hline
Elite Size & 0\\
\hline
\end{tabular}
\end{table}
\end{frame}

\begin{frame}{Terminal Set}
\protect\hypertarget{terminal-set}{}
\begin{tabular}{l|l}
\hline
Terminal & Description\\
\hline
X1 & Relative Compactness\\
\hline
X2 & Surface Area\\
\hline
X3 & Wall Area\\
\hline
X4 & Roof Area\\
\hline
X5 & Overall Height\\
\hline
X6 & Orientation\\
\hline
X7 & Glazing Area\\
\hline
X8 & Glazing Area Distribution\\
\hline
random\_int & Ephemeral Constant (integer)\\
\hline
random\_float & Ephemeral Constant(float)\\
\hline
\end{tabular}
\end{frame}

\begin{frame}{Function Set}
\protect\hypertarget{function-set}{}
\begin{table}

\caption{\label{tab:unnamed-chunk-4}Functions}
\centering
\begin{tabular}[t]{l|r}
\hline
Function & Arity\\
\hline
Addition & 2\\
\hline
Subtraction & 2\\
\hline
Multiplication & 2\\
\hline
Negation & 1\\
\hline
Sine & 1\\
\hline
Cosine & 1\\
\hline
Protected Division & 2\\
\hline
Protected Natural Logarithm & 1\\
\hline
Protected Square Root & 1\\
\hline
\end{tabular}
\end{table}
\end{frame}

\begin{frame}{Examples}
\protect\hypertarget{examples}{}
Model evolved by tournament selection after 100 generations:

\includegraphics{../plots/example_model.png}
\end{frame}
